Data Exploration: IUCN and AquaMaps

### 2015-11-21 data_explore.Rmd
adding map of all species to look at global representation across latitudes

library(ggplot2)
library(maps)
library(readr)
library(raster)
library(data.table)
library(tidyr)
library(dplyr)
library(stringr)

@@

### Species Map Function ###
# This function takes a single species scientific name as input, then grabs all 
# occurrence cells and associated Aquamaps probability and/or IUCN proportional area
# per cell
get_spp_map <- function(species){
  spp_id <- spp_list_base %>%
    filter(sciname == species & is.na(parent_sid)) %>%
    dplyr::select(am_sid, iucn_sid, sciname) %>%
    unique()
  iucn_spp_map <- iucn_spp_cells %>%
    filter(sciname == species) %>%
    group_by(loiczid) %>%        
    summarize(iucn_area = max(prop_area))
  ### The group_by() and summarize() are to collapse cells with overlapped polygons, fixing the duplicate 'by' issue
  am_spp_map   <- am_spp_cells %>%
    filter(am_sid == spp_id$am_sid) %>%
    rename(am_prob = prob)
  spp_map <- full_join(iucn_spp_map, am_spp_map, by = 'loiczid')%>%
    as.data.frame()

@@

A_mdl_clip <- lm(total_area ~ sm_perc, l)
#               Estimate Std. Error t value Pr(>|t|)    
#   (Intercept)  2993939     810560   3.694 0.000227 ***
#   sm_perc       153523      11952  12.844  < 2e-16 ***
#   Multiple R-squared:  0.07955,	Adjusted R-squared:  0.07906 
logA_mdl   <- lm(log_total_area ~ sm_perc, spp_list)
#                Estimate Std. Error t value Pr(>|t|)    
#   (Intercept) 14.775325   0.096446   153.2   <2e-16 ***
#   sm_perc      0.016885   0.001407    12.0   <2e-16 ***
#   Multiple R-squared:  0.0687,	Adjusted R-squared:  0.06823 
```

>>>>

library(rasterVis)
library(ggplot2)
library(maps)
library(readr)
library(raster)
library(data.table)
library(tidyr)
library(dplyr)
library(stringr)

@@

### Species Map Function ###
# This function takes a single species scientific name as input, then grabs all 
# occurrence cells and associated Aquamaps probability and/or IUCN proportional area
# per cell

get_spp_map <- function(species){
  spp_id <- spp_list_base %>%
    filter(sciname == species & is.na(parent_sid)) %>%
    dplyr::select(am_sid, iucn_sid, sciname) %>%
    unique()
  iucn_spp_map <- iucn_spp_cells %>%
    filter(sciname == species) %>%
    group_by(loiczid) %>%        
    summarize(iucn_area = max(prop_area))
  ### The group_by() and summarize() are to collapse cells with overlapped polygons, fixing the duplicate 'by' issue
  am_spp_map   <- am_spp_cells %>%
    filter(am_sid == spp_id$am_sid) %>%
    rename(am_prob = prob)
  spp_map <- full_join(iucn_spp_map, am_spp_map, by = 'loiczid')%>%
    as.data.frame()

@@

A_mdl_clip <- lm(total_area ~ sm_perc, l)
#               Estimate Std. Error t value Pr(>|t|)    
#   (Intercept)  2993939     810560   3.694 0.000227 ***
#   sm_perc       153523      11952  12.844  < 2e-16 ***
#   Multiple R-squared:  0.07955,	Adjusted R-squared:  0.07906 
logA_mdl   <- lm(log_total_area ~ sm_perc, spp_list)
#                Estimate Std. Error t value Pr(>|t|)    
#   (Intercept) 14.775325   0.096446   153.2   <2e-16 ***
#   sm_perc      0.016885   0.001407    12.0   <2e-16 ***
#   Multiple R-squared:  0.0687,	Adjusted R-squared:  0.06823 
```
***

###Global representation of species

By counting the number of unique species are found in each of the half degree cells we can then create a single global raster of number of species.

```{r map, echo=F, warning=F, message=F}

# Number of aquamaps species per cell

am_n_per_cell = filter(am_spp_cells,am_sid %in% unique(spp_list$am_sid))%>%
                  group_by(loiczid)%>%
                   summarize(count.am=n())%>%
                    as.data.frame()

am_sp_map = subs(loiczid_raster, am_n_per_cell[ , c('loiczid', 'count.am')], 
                     by = 'loiczid', 
                     which = 'count.am', 
                     subsWithNA = TRUE)

#number of IUCN species per cell
iucn_n_per_cell = iucn_spp_cells%>%
                   group_by(loiczid)%>%
                    summarize(count.iucn=n())%>%
                    as.data.frame()

iucn_sp_map = subs(loiczid_raster, iucn_n_per_cell[ , c('loiczid', 'count.iucn')], 
                     by = 'loiczid', 
                     which = 'count.iucn', 
                     subsWithNA = TRUE)

#Plot both maps
myTheme=rasterTheme(region=brewer.pal('Blues', n=9))

levelplot(am_sp_map,par.settings=myTheme,main='Aquamaps Dataset',colorkey=list( at=seq(0, 1063, length=15)))

levelplot(iucn_sp_map,par.settings=myTheme,main='IUCN Dataset',colorkey=list(at=seq(0, 1063, length=15)))
```

Now look at all aquamaps species

```{r all aquamaps,eval=F}

am_spp_cells_big = fread(file.path(dir_anx, 'raw/aquamaps_2014/tables/ohi_hcaf_species_native.csv'))
#am_spp_cell_big is missing loiczid info so bring in hcaf to match

hcaf <- read_csv(file.path(dir_anx,'raw/aquamaps_2014/tables/hcaf.csv')) #bring this in just to get loiczid

all_am = am_spp_cells_big%>%
          group_by(loiczid)%>%
          summarize(count=n())%>%
          as.data.frame()

all_am_map = subs(loiczid_raster, all_am[ , c('loiczid', 'count')], 
                     by = 'loiczid', 
                     which = 'count', 
                     subsWithNA = TRUE)

levelplot(all_am_map,par.settings=myTheme,main='Entire Aquamaps Dataset',colorkey=list(at=seq(0, 1063, length=15)))
```
